A Large-Scale Study of the Use of Eval in R

### 2020-08-29 inst/paper/main.tex
sruff
Fix commas

%\input{tag/enrichDataset.tex}
\newcommand{\UndefinedEvalsRnd}{285K\xspace}
\newcommand{\UndefinedEvals}{285,289\xspace}
\newcommand{\PercentUndefinedEval}{3.3\%\xspace}
\newcommand{\UndefinedEvalExpressionsRnd}{262\xspace}
\newcommand{\UndefinedEvalExpressions}{262\xspace}
\newcommand{\UndefinedOutOfCorpusRnd}{0\xspace}
\newcommand{\UndefinedOutOfCorpus}{0\xspace}
\newcommand{\UndefinedOutOfCoprusPercent}{0\%\xspace}
\newcommand{\MaxCallSitesEnrich}{114}
\newcommand{\OneCallSitesEnrich}{61}
\newcommand{\NoCallSitesEnrich}{297}
%\input{tag/patterns.tex}
\newcommand{\PatternLibRnd}{4K\xspace}
\newcommand{\PatternLib}{4,278\xspace}
\newcommand{\PatternLibRatio}{.05\%\xspace}
\newcommand{\PatternLibSourceRnd}{370\xspace}
\newcommand{\PatternLibSource}{370\xspace}
\newcommand{\PatternLibLocalRnd}{100\xspace}
\newcommand{\PatternLibLocal}{100\xspace}
\newcommand{\PatternReadRnd}{59K\xspace}
\newcommand{\PatternRead}{59,776\xspace}
\newcommand{\PatternReadRatio}{.7\%\xspace}
\newcommand{\PatternReadSlotAccessRatio}{88\%\xspace}
\newcommand{\PatternReadDatatableRatio}{27\%\xspace}
\newcommand{\PatternAssignRnd}{68K\xspace}
\newcommand{\PatternAssign}{68,754\xspace}
\newcommand{\PatternAssignRatio}{.9\%\xspace}
\newcommand{\PatternAssignArrowRatio}{18\%\xspace}
\newcommand{\PatternAssignArrowMagrittrRatio}{59\%\xspace}
\newcommand{\PatternAssignSuperarrowRatio}{81\%\xspace}
\newcommand{\PatternAssignSuperarrowPlyrRatio}{99\%\xspace}
\newcommand{\PatternManualPackagesRnd}{20\xspace}
\newcommand{\PatternManualPackages}{20\xspace}
\newcommand{\PatternManualCallsitesRnd}{161\xspace}
\newcommand{\PatternManualCallsites}{161\xspace}

@@

\newcommand{\eval}{\c{eval}}
\newcommand{\evals}{{\sf eval}s\xspace}
\newcommand{\parse}{\c{parse}}
\newcommand{\source}{\c{source}}
\newcommand{\local}{\c{local}}
\newcommand{\unlockBinding}{\c{unlockBinding}}
\newcommand{\substitute}{\c{substitute}}
\newcommand{\datatable}{\c{data.table}}
\newcommand{\mlogit}{\c{mlogit}}
\newcommand{\mboost}{\c{mboost}}
\newcommand{\metafor}{\c{metafor}}
\newcommand{\lavaan}{\c{lavaan}}
\newcommand{\mclust}{\c{mclust}}
\newcommand{\gamlss}{\c{gamlss}}
\newcommand{\ggproto}{\c{ggproto}}
\newcommand{\ggplot}{\c{ggplot2}}
\newcommand{\base}{\c{base}}
\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}
\newcommand{\evil}{\emph{evil}\xspace}
\newcommand{\instrumentr}{\emph{instrumentr}\xspace}
\newcommand{\rdyntrace}{\emph{R-dyntrace}\xspace}
\newcommand{\covr}{\emph{covr}\xspace}
\newcommand{\runr}{\emph{runr}\xspace}

@@

Unlike CRAN, Kaggle is not a curated repository and therefore there are no
guarantees about the quality of the code. After downloading all of the
\CorpusKaggle kernels and extracting the R code from the various
formats,\footnote{We use {\sf rmarkdown} to convert from notebooks to R.} we
found that \CorpusDuplicatedKaggle were whole-file duplicates. From
\CorpusRunnableKaggle kernels, \CorpusFailedKaggle failed to execute. Next
to various runtime exceptions, common problems were parse errors and
misspelled package names.  The final set contains \CorpusFinishedKaggle
kernels with \summaryrnd{CorpusFinishedKaggleCode} lines of R code.

@@

This section presents an overview of the usage of \eval in our corpus.
Richards et al. reported 82\% of JavaScript-enabled web pages using
\eval~\cite{ecoop11}. The case for R is more striking with 100\% of the
programs call the \eval function.\footnote{We combine calls to {\sf eval},
  {\sf eval.parent}, {\sf evalq} and {\sf local}.}  This high number is
explained by the fact that the R implementation, its core packages, use
\eval extensively. We differentiate uses based on their origin. A {\bf core}
\eval is one that occurs in the code of a core package.  A {\bf dependent}
\eval occurs in a package imported by the target package. An {\bf own} \eval
is one that occurs in the source code of the target package.

@@

Table~\ref{table:site-package-summary} shows the number of call sites in
CRAN packages. \OneCallSitesEnrich packages have a single call site, and one
package, \emph{metafor}, has \MaxCallSitesEnrich call sites of
\eval. \NoCallSitesEnrich packages have no calls to \eval. The core packages
have \NbCoreEvalCallSites call sites of \eval and Kaggle has none. These
numbers are under-approximations as \eval may be called reflectively,
through aliases, or passed to a higher-order function.

@@

In this section we turn our attention to the expression passed for
evaluation to \eval. The \eval function accepts all types of R objects as
inputs, but in the majority of cases (\AllExpressionInputEvalCallPerc), it
is called with expressions. The remaining \AllValueInputEvalCallPerc are
values, \ie, they evaluate to themselves.
%
Passing a value into an eval is effectively an no-op so we look closely at
the expression types the call sites pass to eval across all the runs.
\AllValueInputEvalSitePerc of all call sites always pass values,
\AllExpressionInputEvalSitePerc pass expressions and a very large
proportion, \AllPolymorphicInputEvalSitePerc, pass both expressions and
values as inputs across all program runs. We inspected the eval callsites
that accept only values or both values and expression as inputs. The top
three of these call sites contribute to over ~80\% of all calls to
evals. The first one is inside \c{match.arg} from core. Given a vector of
names as strings, this function finds the default values for parameters with
those names in the caller function. Since the default value for a function
parameter can be a complex expression or a benign value such as \c{NULL},
\c{match.arg} evaluates it through a call to \eval. This makes the \eval
inside \c{match.arg} ``polymorhpic''. The second such \eval call originates
from the popular \ggplot package of R which is a DSL for plotting. \ggplot
has it's own object system, \ggproto, which uses \eval inside the \ggproto
function to determine the super-class from which it inherits. The default
value of the superclass is \c{NULL} but in general could be an arbitrarily
complex expression. The third \eval call is in core R \c{str} function. This
function is used to print a summary of R objects. It internally maps \eval
function on the bindings of an environment which results in inputs of
multiples types being passed to \eval.

@@

\begin{table}[!h] \centering
\begin{tabular}{@{}l|rr@{}} \hline
Expression & Eval Call &  \% \\\hline
\c{c("auto", "shell", "radix")} & 198,7105 & 29\%\\
\c{c("auto", "shell", "quick", "radix")} & 1,593,169  & 23\%\\
\c{\{info <- loadingNamespaceInfo(...} & 1,008,632 &       14\%\\
\c{c("onLoad", "attach", "detach", "onUnload")}   & 470,566 &      6.9\%\\
\c{c("append", "prepend", "replace")} &              261,587&       3.9\% \\
\c{c("left", "right", "centre", "none")} & 162,086     & 2.4\%\\
\c{c("no", "ifany", "always")}   &                71,580 &       1.1\%\\
\c{c("pearson", "kendall", "spearman")}  & 72,962 &      1.1\%\\
\c{NULL}& 75330  &      1.1\% \\
\c{Symbol}&                 6,6279&       1\%\\\hline
\end{tabular}
\caption{Top ten eval calls in Core}\label{B}
\end{table}

@@

\begin{table}[ht]%\resizebox{\columnwidth}{!}{%
  \centering
\begin{tabular}{r|r|r|r|r}\hline
\multirow{2}{*}{Pattern} & \multicolumn{2}{c|}{Core} & \multicolumn{2}{c}{Packages} \\
                         & Calls & Ratio            & Calls & Ratio \\\hline
\input{tag/table-js-patterns-in-r.tex}
\end{tabular}
\caption{JavaScript evals patterns in R}
\label{table:js-pattens-in-r}
\end{table}

\noindent \emph{Library}, \emph{Typeof} and \emph{Try} are all barely used
mostly because the language provides dedicated features for it. The few calls in
Library pattern come mostly from file sourcing functions (\ie \c{source} and \c{sys.source}).
%
\noindent \emph{Read} pattern in JavaScript represented mostly an incorrect
access to an object's property or local variable both replaceable by
JavaScript core functions. In the case of R there are \PatternReadRnd read
\eval calls (\PatternReadRatio). They come from calls like
\c{subset(airquality, select = Temp)} where the variable \c{Temp} will be
evaluated in the \c{airquality} data frame.
%
\noindent \emph{Assign} category comprises of assignments to variables
within an \eval call. We have seen \PatternAssignRnd such calls
(\PatternAssignRatio).  \PatternAssignArrowRatio uses the simple assignment
(\c{<-}). The majority of cases comes from the \c{magrittr}'s \c{\%<>\%}
pipe operator (\PatternAssignArrowMagrittrRatio) which forwards an object to
a function on rhs and update lhs with the result. Far more calls used the
super assignment (\c{<<-}) operator, but almost all come from the \c{plyr}
package which provides a convenient API for common data manipulation
problems.
%
\noindent \emph{Call} and \emph{Other} together represents the vast majority
of eval calls in R. The \emph{Other} pattern represents all the calls where
eval was called with a value instead of with an expression. The \emph{Call}
pattern contains all the calls that did not fall in the previous
categories. Without a context it is usually difficult to further categorize
the intent of these calls.  Therefore we have randomly chosen
\PatternManualPackages packages and manually looked at
\PatternManualCallsites call sites from which we have observed the following
categories:

>>>>

%\input{tag/enrichDataset.tex}
\newcommand{\UndefinedEvalsRnd}{285K\xspace}
\newcommand{\UndefinedEvals}{285,289\xspace}
\newcommand{\PercentUndefinedEval}{3.3\%\xspace}
\newcommand{\UndefinedEvalExpressionsRnd}{262\xspace}
\newcommand{\UndefinedEvalExpressions}{262\xspace}
\newcommand{\UndefinedOutOfCorpusRnd}{0\xspace}
\newcommand{\UndefinedOutOfCorpus}{0\xspace}
\newcommand{\UndefinedOutOfCoprusPercent}{0\%\xspace}
\newcommand{\MaxCallSitesEnrich}{114\xspace}
\newcommand{\OneCallSitesEnrich}{61\xspace}
\newcommand{\NoCallSitesEnrich}{297\xspace}
%\input{tag/patterns.tex}
\newcommand{\PatternLibRnd}{4K\xspace}
\newcommand{\PatternLib}{4,278\xspace}
\newcommand{\PatternLibRatio}{.05\%\xspace}
\newcommand{\PatternLibSourceRnd}{370\xspace}
\newcommand{\PatternLibSource}{370\xspace}
\newcommand{\PatternLibLocalRnd}{100\xspace}
\newcommand{\PatternLibLocal}{100\xspace}
\newcommand{\PatternReadRnd}{59K\xspace}
\newcommand{\PatternRead}{59,776\xspace}
\newcommand{\PatternReadRatio}{.7\%\xspace}
\newcommand{\PatternReadSlotAccessRatio}{88\%\xspace}
\newcommand{\PatternReadDatatableRatio}{27\%\xspace}
\newcommand{\PatternAssignRnd}{68K\xspace}
\newcommand{\PatternAssign}{68,754\xspace}
\newcommand{\PatternAssignRatio}{.9\%\xspace}
\newcommand{\PatternAssignArrowRatio}{18\%\xspace}
\newcommand{\PatternAssignArrowMagrittrRatio}{59\%\xspace}
\newcommand{\PatternAssignSuperarrowRatio}{81\%\xspace}
\newcommand{\PatternAssignSuperarrowPlyrRatio}{99\%\xspace}
\newcommand{\PatternManualPackagesRnd}{20\xspace}
\newcommand{\PatternManualPackages}{20\xspace}
\newcommand{\PatternManualCallsitesRnd}{161\xspace}
\newcommand{\PatternManualCallsites}{161\xspace}

@@

\newcommand{\eval}{\texttt{eval}}
\newcommand{\evals}{{\sf eval}s\xspace}
\newcommand{\parse}{\c{parse}}
\newcommand{\source}{\c{source}}
\newcommand{\local}{\c{local}}
\newcommand{\unlockBinding}{\c{unlockBinding}}
\newcommand{\substitute}{\c{substitute}}
\newcommand{\datatable}{\c{data.table}}
\newcommand{\mlogit}{\c{mlogit}}
\newcommand{\mboost}{\c{mboost}}
\newcommand{\metafor}{\c{metafor}}
\newcommand{\lavaan}{\c{lavaan}}
\newcommand{\mclust}{\c{mclust}}
\newcommand{\gamlss}{\c{gamlss}}
\newcommand{\ggproto}{\c{ggproto}}
\newcommand{\ggplot}{\c{ggplot2}}
\newcommand{\base}{\c{base}}
\renewcommand{\c}[1]{\lstinline{#1}\xspace}
\newcommand{\miss}[1]{{\textcolor{red}{#1}}\xspace}
\newcommand{\evil}{\emph{evil}\xspace}
\newcommand{\instrumentr}{\emph{instrumentr}\xspace}
\newcommand{\rdyntrace}{\emph{R-dyntrace}\xspace}
\newcommand{\covr}{\emph{covr}\xspace}
\newcommand{\runr}{\emph{runr}\xspace}

@@

Unlike CRAN, Kaggle is not a curated repository and therefore there are no
guarantees about the quality of the code. After downloading all of the
\CorpusKaggle kernels and extracting the R code from the various
formats,\footnote{We use {\sf rmarkdown} to convert from notebooks to R.} we
found that \CorpusDuplicatedKaggle were whole-file duplicates. From
\CorpusRunnableKaggle kernels, \CorpusFailedKaggle failed to execute. Next
to various runtime exceptions, common problems were parse errors and
misspelled package names.  The final set contains \CorpusFinishedKaggle
kernels with \CorpusFinishedKaggleCodeRnd lines of R code.

@@

This section presents an overview of the usage of \eval in our corpus. Richards
et al. reported 82\% of JavaScript-enabled web pages using \eval~\cite{ecoop11}.
The case for R is more striking with 100\% of the programs call the \eval
function.\footnote{We combine calls to {\sf eval}, {\sf eval.parent}, {\sf
    evalq} and {\sf local}.} This high number is explained by the fact that the
R implementation, its core packages, use \eval extensively. We differentiate
three types of eval based on their origin:
%
\begin{compactitem}[$-$]
  \item a {\bf core eval} is one that occurs in the code of a core package,
  \item a {\bf dependent eval} occurs in a package imported by the target
  package, and
  \item an {\bf own eval} is one that occurs in the source code of the target
  package.
\end{compactitem}

@@

Table~\ref{table:site-package-summary} shows the number of call sites in
CRAN packages. \OneCallSitesEnrich{} packages have a single call site, and one
package, \emph{metafor}, has \MaxCallSitesEnrich{} call sites of
\eval. \NoCallSitesEnrich{} packages have no calls to \eval. The core packages
have \NbCoreEvalCallSites{} call sites of \eval and Kaggle has none. These
numbers are under-approximations as \eval may be called reflectively,
through aliases, or passed to a higher-order function.

@@

In this section we turn our attention to the expression passed for
evaluation to \eval. The \eval function accepts all types of R objects as
inputs, but in the majority of cases (\AllExpressionInputEvalCallPerc), it
is called with expressions. The remaining \AllValueInputEvalCallPerc are
values, \ie, they evaluate to themselves.
%
Passing a value into an eval is effectively an no-op so we look closely at the
expression types the call sites pass to eval across all the runs.
\AllValueInputEvalSitePerc of all call sites always pass values,
\AllExpressionInputEvalSitePerc pass expressions and a very large proportion,
\AllPolymorphicInputEvalSitePerc, pass both expressions and values as inputs
across all program runs. We inspected the eval callsites that accept only values
or both values and expression as inputs. The top three of these call sites
contribute to over ~80\% of all calls to evals. The first one is inside
\c{match.arg} from core. Given a vector of names as strings, this function finds
the default values for parameters with those names in the caller function. Since
the default value for a function parameter can be a complex expression or a
benign value such as \c{NULL}, the \c{match.arg} function evaluates it through a
call to \eval. This makes the \eval inside \c{match.arg} ``polymorhpic''. The
second such \eval call originates from \ggplot, a popular R package for making
plots. \ggplot has it's own object system, \ggproto, which uses \eval inside the
\ggproto function to determine the super-class from which it inherits (the
default value is \c{NULL} but in general it could be an arbitrarily expression).
The third \eval call is in core R \c{str} function used to print a summary of R
objects. It internally maps \eval function on the bindings of an environment
which results in inputs of multiples types being passed to \eval.

@@

\begin{table}[!h] \centering
\begin{tabular}{@{}l|rr@{}} \hline
Expression & Eval Call &  \% \\\hline
\c{c("auto", "shell", "radix")} & 1,987,105 & 29\%\\
\c{c("auto", "shell", "quick", "radix")} & 1,593,169  & 23\%\\
\c{\{info <- loadingNamespaceInfo(...} & 1,008,632 &       14\%\\
\c{c("onLoad", "attach", "detach", "onUnload")}   & 470,566 &      6.9\%\\
\c{c("append", "prepend", "replace")} &              261,587&       3.9\% \\
\c{c("left", "right", "centre", "none")} & 162,086     & 2.4\%\\
\c{c("no", "ifany", "always")}   &                71,580 &       1.1\%\\
\c{c("pearson", "kendall", "spearman")}  & 72,962 &      1.1\%\\
\c{NULL}& 75,330  &      1.1\% \\
\c{Symbol}&                 66,279&       1\%\\\hline
\end{tabular}
\caption{Top ten eval calls in Core}\label{B}
\end{table}

@@

\begin{table}[ht]%\resizebox{\columnwidth}{!}{%
  \centering
\begin{tabular}{r|r|r|r|r}\hline
\multirow{2}{*}{Pattern} & \multicolumn{2}{c|}{Core} & \multicolumn{2}{c}{Packages} \\
                         & Calls & Ratio            & Calls & Ratio \\\hline
\input{tag/table-js-patterns-in-r.tex}
\end{tabular}
\caption{JavaScript \eval Patterns in R}
\label{table:js-pattens-in-r}
\end{table}

\noindent \emph{Library}, \emph{Typeof} and \emph{Try} are all barely used
mostly because the language provides dedicated features for it. The few calls in
Library pattern come mostly from file loading functions such as \c{source} and
\c{sys.source}.
%
\noindent \emph{Read} pattern in JavaScript represented mostly an incorrect

access to an object's property or local variable both replaceable by
JavaScript core functions. In the case of R there are \PatternReadRnd read
\eval calls (\PatternReadRatio). They come from calls like
\c{subset(airquality, select = Temp)} where the variable \c{Temp} will be
evaluated in the \c{airquality} data frame.

%
\noindent \emph{Assign} category comprises of assignments to variables
within an \eval call. We have seen \PatternAssignRnd such calls
(\PatternAssignRatio).  \PatternAssignArrowRatio uses the simple assignment
(\c{<-}). The majority of cases comes from the \c{magrittr}'s \c{\%<>\%}
pipe operator (\PatternAssignArrowMagrittrRatio) which forwards an object to
a function on rhs and update lhs with the result. Far more calls used the
super assignment (\c{<<-}) operator, but almost all come from the \c{plyr}
package which provides a convenient API for common data manipulation
problems.
%
\noindent \emph{Call} and \emph{Other} together represents the vast majority
of eval calls in R. The \emph{Other} pattern represents all the calls where
eval was called with a value instead of with an expression. The \emph{Call}
pattern contains all the calls that did not fall in the previous
categories. Without a context it is usually difficult to further categorize
the intent of these calls.  Therefore we have randomly chosen
\PatternManualPackages packages and manually looked at
\PatternManualCallsites call sites from which we have observed the following
categories:
